---
title: 'Class 2 Assignment'
author: 'Nate Lindley'
date: '9/1/2022'
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
---

# Dataset 1 (beaches.csv)

```{r}
library(tidyr)
library(dplyr)
beaches <- read.csv('~/DS-3001/data/beaches.csv')
head(beaches)
# summary(beaches)
```
## Quesiton 1

### Question

What temperature does the most rainfall happen at?

### Psuedocode

1) mutate column for temp rounded to nearest 5
2) group by temp (rounded to nearest 5)
3) average rainfall
4) remove NA's (set to 0 since that's what they represent)
5) sort descending

### Single Line Code

```{r}
beaches1 = mutate(beaches, temp5 = 5* round(beaches$temperature/5))

beaches2 = group_by(beaches1, temp5) %>%
  summarize(avg_rainfall = mean(rainfall))

beaches3 = na.omit(beaches2)

beaches4 = arrange(beaches3, -avg_rainfall)

head(beaches4)
```

### Function

```{r}
avg_rainfall = function(df) {
  mutate(df, temp5 = 5* round(beaches$temperature/5)) %>%
    group_by(temp5) %>%
    summarize(avg_rainfall = mean(rainfall)) %>%
    na.omit() %>%
    arrange(-avg_rainfall)
}

beaches_func = avg_rainfall(beaches)
head(beaches_func)
```

## Question 2

### Question

For this one I removed some columns of the dataset, the split up date columns,
and will go through the steps to recreate them.

```{r}
beaches_stripped = select(beaches,-c('day', 'month', 'year'))
```

### Psuedocode

1) split after the year
2) remove the dash before the month digits
3) split after month
4) remove dash before day digits
5) cast all three columns to numeric

### Single Line Code

```{r}
beach1 = separate(beaches_stripped, date, c('Year', 'mm-dd'), sep=4, remove=FALSE)

beach2 = separate(beach1, 'mm-dd', c('_', 'mm-dd'), sep=1)

beach3 = select(beach2, -'_')

beach4 = separate(beach3, 'mm-dd', c('Month','-dd'), sep=2)

beach5 = separate(beach4, '-dd', c('_', 'Day'), sep=1)

beach6 = select(beach5, -'_')

num_cols = c('Year','Month','Day')

beach6[num_cols] = sapply(beach6[num_cols], as.numeric)
```

### Function

```{r}
recreate = function(df) {
  new_df = separate(df, date, c('Year', 'mm-dd'), sep=4, remove=FALSE) %>%
    separate('mm-dd', c('_', 'mm-dd'), sep=1) %>%
    select(-'_') %>%
    separate('mm-dd', c('Month','-dd'), sep=2) %>% 
    separate('-dd', c('_', 'Day'), sep=1) %>%
    select(-'_')

  num_cols = c('Year','Month','Day')
  new_df[num_cols] = sapply(new_df[num_cols], as.numeric)
  new_df
}

new_beach = recreate(beaches_stripped)
head(new_beach)
```


# Dataset 2: NFL Dataset

```{r}
nfl <- read.csv("~/DS-3001/data/nfl2008_fga.csv")
head(nfl)
```

## Quesiton 1

### Question

Which kicker had the highest fg make percentage in 2008?

### Psuedocode

1) group_by kicker name and sum up makes, blocks, and misses
2) mutate additional column with make %
3) sort in descending order

### Single Line Code

```{r}

nfl1 = group_by(nfl, name) %>%
  summarize(GOOD = sum(GOOD == 1), BAD = sum(Missed == 1 | Blocked == 1))

nfl2 = mutate(nfl1, pct = GOOD/(GOOD+BAD))

nfl3 = arrange(nfl2, -pct)

head(nfl3)

```

### Function

```{r}

best_kicker = function(df) {
  group_by(df, name)%>%
    summarize(GOOD = sum(GOOD == 1), BAD = sum(Missed == 1 | Blocked == 1))%>%
    mutate(pct = GOOD/(GOOD+BAD))%>%
    arrange(-pct)
}

nfl_func1 = best_kicker(nfl)
head(nfl_func1)

```

## Question 2

### Question

How does fg attempt distance affect make percentage?

### Psuedocode

1) mutate column for distance10 (rounding to nearest 10)
2) group by distance10, counting makes and misses
3) mutate column for make %

### Single Line Code

```{r}
nfl_a = mutate(nfl, distance10 = 10*round(nfl$distance/10))

nfl_b = group_by(nfl_a, distance10) %>%
  summarize(GOOD = sum(GOOD == 1), BAD = sum(Missed == 1 | Blocked == 1))

nfl_c = mutate(nfl_b, pct = GOOD/(GOOD+BAD))

head(nfl_c)
```

### Function

```{r}
distance_summary = function(df) {
  mutate(df, distance10 = 10*round(nfl$distance/10)) %>%
    group_by(distance10) %>%
    summarize(GOOD = sum(GOOD == 1), BAD = sum(Missed == 1 | Blocked == 1)) %>%
    mutate(pct = GOOD/(GOOD+BAD))
}

dsumm = distance_summary(nfl)
head(dsumm)
```