---
title: "DT Lab"
author: "Nathan Lindley"
date: "October 20, 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(mice)
library(dplyr)
library(caret)
library(rpart.plot)
library(ROCR)
#install.packages("MLmetrics")
library(MLmetrics)
library(mltools)
library(data.table)
```

Congrats! You just graduated UVA's MSDS program and got a job working at a movie studio in Hollywood. 

Your boss is the head of the studio and wants to know if they can gain a competitive advantage by predicting new movies that might get high imdb scores (movie rating). 

You would like to be able to explain the model to the mere mortals 
but need a fairly robust and flexible approach so you've 
chosen to use decision trees to get started. 

In doing so, similar to great data scientists of the past 
you remembered the excellent education provided 
to you at UVA in a undergrad data science course and have outlined 
20ish steps that will need to be undertaken to complete this task. As always, you will need to make sure to #comment your work heavily. 


 Footnotes: 
-	You can add or combine steps if needed
-	Also, remember to try several methods during evaluation and always be 
mindful of how the model will be used in practice.
- Make sure all your variables are the correct type (factor, character, numeric, etc.)


```{r}
#1 Load the data and ensure the labels are correct. restate the question and comment on evaluation metrics that you might pay attention to.  
df <- read_csv("~/DS-3001/data/movie_metadata.csv")
colnames(df)

# Question: How can we predict which movies will receive a high IMDB score?
# Evaluation Metrics: Sensitivity, loss of movie revenue from a missed hit would be disastrous
```

```{r}
#2 Ensure all the variables are classified correctly including the target variable and collapse factor variables as needed. 
drop_cols <- c(2,7,10,11,12,15,17,18)
df <- df[-drop_cols]
str(df)
```

```{r}
#3 Check for missing variables and correct as needed.  
# mice::md.pattern(df)

df <- df[complete.cases(df),]

med <- median(df$imdb_score)

df <- df%>%
  mutate(aboveMedian = ifelse(df$imdb_score > med, 'True', 'False'))

df$content_rating <- make.names(df$content_rating)
df$country <- make.names(df$country)
df$color <- make.names(df$color)
df$language <- make.names(df$language)
df$aboveMedian <- make.names(df$aboveMedian)
# df = lapply(df, make.names)

factor_columns <- c(1, 12, 13, 14, 21)
df[,factor_columns] <- lapply(df[,factor_columns], as_factor)
colnames(df) <- make.names(colnames(df))
df <- df[,-18]
```

```{r}
#4 Guess what, you don't need to standardize the data,because DTs don't require this to be done, they make local greedy decisions...keeps getting easier, go to the next step
```

```{r}
#5 Determine the baserate or prevalence for the classifier, what does this number mean? (Need to cut the target appropriately)

table(df$aboveMedian)
# 0: 2024, 1: 1746, Total: 3770
# I'm assuming these are different because there are a significant number of ratings
# that are equal to the median.
prevalence = 1746/3770
prevalence # 46.3%
```

```{r}
#6 Split your data into test, tune, and train. (80/10/10)
split_index <- createDataPartition(df$aboveMedian, p=.8,
                                   list=F,
                                   times=1)

train <- df[split_index,]
dim(train)

test <- df[-split_index,]
dim(test)

tuneTestSplit <- createDataPartition(test$aboveMedian, p=.5,
                                     list=F,
                                     times=1)

tune <- test[tuneTestSplit,]
dim(tune)

test <- test[tuneTestSplit,]
dim(test)
```

```{r}
#7 Build your model using the training data, rpart2, and repeated cross validation as reviewed in class with the caret package.
fitControl <- trainControl(method = "repeatedcv",
                          number = 10,
                          repeats = 5,
                          classProbs = T,
                          summaryFunction = twoClassSummary) 

features <- train[,-20]
target <- train$aboveMedian

set.seed(1999)
myDT <- train(x=features,
                y=target,
                method="rpart2",
                trControl=fitControl,
                metric="ROC")
```

```{r}
#8 View the results, comment on how the model performed and which variables appear to be contributing the most (variable importance)  
myDT
plot(myDT)
varImp(myDT)
```

```{r}
#9 Plot the output of the model to see the tree visually, using rpart.plot 
rpart.plot(myDT$finalModel, type=4)
```

```{r}
#10 Use the tune set and the predict function with your model to the 
# predict the target variable, making sure to produce probabilities.
prob_pred <- predict(myDT, tune, type='prob')
prob_pred
```

```{r}
#11 Use the the confusion matrix function on your predictions to check a variety of metrics and comment on the metric that might be best for this type of analysis given your question.  
predictandCM<- function(model,data,modeltype,ref)
{
  #model using, data going into the model, and output type for predict function
  pred <-predict(model,data,type=modeltype)
  confusionMatrix(pred, reference=ref, positive = 'True')
}

predictandCM(myDT,tune,"raw",tune$aboveMedian)
# It looks like the sensitivity is strong on this model, significantly stronger than 
# basic guessing. Noting how we had a 46% prevalence and a 73% sensitivity.
```

```{r}
#12 With the percentages you generated in step 10,select several different threshold levels using the threshold function we created and interpret the results. What patterns do you notice, did the evaluation metrics change? 
densityplot(prob_pred$True)
# Going solely off of this density plot it appears that there is most likely little to be improved by adjusting
# the threshold but we'll still try some

adjust_thres <- function(x, y, z) {
  #x=pred_probablities, y=threshold, z=tune_outcome
  thres <- as.factor(ifelse(x > y, "True","False"))
  confusionMatrix(thres, z, positive = "True", dnn=c("Prediction", "Actual"), mode = "everything")
}

adjust_thres(prob_pred$True, y=.3, tune$aboveMedian) #sensitivity = 83.4%
adjust_thres(prob_pred$True, y=.4, tune$aboveMedian) #sensitivity = 70.9%
adjust_thres(prob_pred$True, y=.5, tune$aboveMedian) #sensitivity = 70.9%
adjust_thres(prob_pred$True, y=.6, tune$aboveMedian) #sensitivity = 65.7%

# It looks like .4 and .5  give the same classifications suggesting we have no percentages between 0.4 and 0.5
```

```{r}
#13 Based on your understanding of the model and data adjust the hyper-parameter via the built in train control function in caret or build and try new features, does the model quality improve or not? If so how and why, if not, why not?
tree.grid <- expand.grid(maxdepth=c(5,7,9,11))

set.seed(1999)
myDT1 <- train(x=features,
                y=target,
                method="rpart2",
                trControl=fitControl,
                tuneGrid=tree.grid,
                metric="ROC")

predictandCM(myDT1,tune,"raw",tune$aboveMedian)
# increased sensitivity from 70.9% at .5 threshold to 73.7%
```

```{r}
#14 Once you are confident that your model is not improving, via changes implemented on the training set and evaluated on the the tune set, predict with the test set and report final evaluation of the model. Discuss the output in comparison with the previous evaluations.  
predictandCM(myDT1,test,"raw",test$aboveMedian)
# On the test dataset I'm getting the exact same specificity and all other metrics
```

```{r}
#15 Summarize what you learned along the way and make recommendations on how this could be used moving forward, being careful not to over promise. 

# Based on my model it appears that with respect to features that can be influenced by the director and 
# producers of a movie the most influential one is run time, specifically over 1 hour and 51 minutes.
# With this in mind I'd recommend making a movie that is longer than 1 hour and 51 minutes as well as 
# one that has a budget below $31 million because it appears that these movies do better.
```

```{r}
#16 What was the most interesting or hardest part of this process and what questions do you still have? 

# I find it weird and interesting that the movies with a budget of below $31 million generally do better than
# the movies with a budget of over $31 million.
```


